Selective electroporation of the viral envelope

### 2021-03-22 documents/pure_electroporation.tex
electroporation

\title{Selective electroporation of the viral envelope}
\author{\small{Daniel Correia\orcidlink{0000-0002-9353-0216} (therobotist@gmail.com), Ontario, Canada.}}
\date{March 2021}

@@

\begin{abstract}
	\noindent Extreme electric fields drive transmembrane potentials, leading to pores through membranes; this phenomenon is commonly harnessed for transfection into the cell in the laboratory, and similar techniques have recently been shown to be effective in the clinic for ablating tumors.\\

	\noindent As pulse duration is reduced from the microsecond to nanosecond regime, effects targeting smaller intracellular structures are expected and observed. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the cell membrane intact.

	Recently-developed gigawatt pulse systems may offer a possible means of noninvasively propagating such pulses into deep tissues, such as within lungs. 
\end{abstract}

As a followup to.

@@

Pulse count and inter-pulse timing is another free variable. Joshi et al \cite{Selfconsistent2001} model that a carefully timed second pulse can have greater effects than a single pulse, in agreement with experiment.

@@

\section*{Basic scaling laws}

Some scaling laws for membrane charging effects due to short (sub Tc) pulses are given by Schoenbach \cite{Bioelectric2007}.

The membrane charging time constant Cole single-shell model \cite{Ultrashort2004} - simplified for the low-density case -

@@

where $\rho_{medium}$ and $\rho_{internal}$ are the resistivities of the medium and internal fluid (the cytoplasm for cells, and assumed to be the 'core' for viruses). \cite{Bioelectric2007} eq. 15,

@@

pore size vs time? same size pore in a smal lparticle

@@

that is, all else being equal, the induced transmembrane voltage appears to be invariant with raw particle diameter - contrary to Rjif's more detailed analysis.

@@

In electrorotation, it is usually found that the viral shell has a permittivity of about 60 to 80, similar to the nuclear membrane (60 to 100) but massively unlike the cell membrane ($\epsilon_r\approx 10$).

@@

However, if a pulse train with a pulse duration shorter than the cell time constant, an intensity too low for, and a repetition rate of some 3 us, (200 khz),

>>>>

\title{Selective electroporation of the viral envelope}
\author{\small{Daniel Correia\orcidlink{0000-0002-9353-0216} (therobotist@gmail.com), Ontario, Canada.}}
\date{March 21 2021}

@@

\begin{abstract}
	\noindent The phenomenon of electroporation occurs when extreme applied electric fields in the MV/m range drive transmembrane potentials, leading to pores through membranes; this is commonly harnessed for transfection into the cell in the laboratory, and has recently been shown to be safe and effective in the clinic for ablating tumors.\\

	\noindent As pulse duration is reduced from the microsecond to nanosecond regime, effects targeting smaller intracellular structures are expected and observed. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the plasma membrane intact. \\

	Dielectric properties apparently due to the

	\noindent Recently-developed gigawatt pulse systems may offer a possible means of noninvasively propagating such pulses into deep tissues, such as within lungs. 
\end{abstract}

As a followup to \cite{notes2021}

The same electrostatic effects on lipid bilayers have been investigated for decades. Many extremely fascinating hydrodynamic and parametric

@@

Pulse count and inter-pulse timing is another free variable. Joshi et al \cite{Selfconsistent2001} model that a carefully timed second pulse can have greater effects than a single pulse, in agreement with experiment on E. coli.

@@

This seems too simple.

\subsection*{Basic scaling laws}

Some extremely simplified scaling laws for membrane charging effects due to short (sub Tc) pulses are given by Schoenbach \cite{Bioelectric2007}. These are based primarily on an equivalent circuit model.

The membrane charging time constant from a single-shell model \cite{Ultrashort2004} - simplified for the low-density case

@@

where $\rho_{medium}$ and $\rho_{internal}$ are the resistivities of the medium and internal fluid (the cytoplasm for cells, and assumed to be the 'core' for viruses). Electroporation is a highly nonlinear effect, and this . \cite{Bioelectric2007} eq. 15,

@@

@@

that is, all else being equal, the induced transmembrane voltage appears to be invariant with raw particle diameter - mildly contrary to Rjif's more detailed analysis.

@@

In electrorotation, it is usually found that the viral shell has a permittivity of about 60 to 80, similar to the nuclear membrane (60 to 100) but massively unlike the cell membrane ($\epsilon_r\approx 10$). This is very remarkable, since the lipid bilayer itself is somewhat similar in composition to the host cell; envelope, spike and membrane proteins may account.

@@

However, if a monopolar pulse train with a pulse duration shorter than the cell time constant, an intensity too low for, and a repetition rate of some 3 us, (200 khz),
